---
title: "Deutschland Tour 2024"
description: |
  Exploring the route of this years 'Deutschland Tour' with R
author:
  - name: Julian During
date: "2024-08-07"
output: distill::distill_article
creative_commons: CC BY
repository_url: https://github.com/duju211/deutschland_tour
base_url: https://www.datannery.com/posts/deutschland_tour/
bibliography: biblio.bib
editor_options: 
  chunk_output_type: console
params:
  interactive: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = !params$interactive)

source("libraries.R")

df_manifest <- tar_manifest() |>
  mutate(command = str_glue("{name} <- {command}"))
```

# Idea

The "Deutschland Tour" is a big (road) bike race in Germany.
By far not as big as the "Tour de France", but maybe it will get there :).

This years "Deutschland Tour" is a special one. It passes near my hometown and
therefore I am wondering where there is a good spot for spectators.
For me these key points
are important:

* Not too far away
* Nice mountain climb (otherwise the riders pass very quickly)
* Nearby drinks are welcome :)

To answer this question, I will use my R skills to import and visualise the
data.

# Reproducibility

In this analysis the following libraries are used:

```{r, file="libraries.R", eval=FALSE}
```

If you want to reproduce this analysis, you have to perform the following steps:

* Clone the repository (see footnote)
* Run `renv::restore()` [@cit_renv]
* Run `targets::tar_make()` [@cit_targets]

Alternatively, you could run this analysis by copying and executing chunk by
chunk in your R session (installing the above mentioned packages manually). 

# Data

At first define where gpx files can be read from:

```{r gpx_url, code=df_manifest$command[df_manifest$name == "gpx_url"]}
```

Define helper function that reads in stage data. To do this first download
the gpx file (using 'httr2' [@cit_httr2]) and save it to an temporary file
(using 'fs' [@cit_fs]).

The resulting file is a xml file. Read it in 
(using the 'xml2' package [@cit_xml2])
and get all the elements that represent "trackpoints". 
Extract spatial information from these "trackpoints" using various
helper functions (see documentation links):

```{r, file="R/stage.R"}
```

Apply the above mentioned function to all urls resulting in one final data
frame:

```{r, code=df_manifest$command[df_manifest$name == "df_stages"]}
```

# Analysis

Turn data frame into a sf [@cit_sf] object:

```{r, code=df_manifest$command[df_manifest$name == "sf_stages"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(sf_stages)
```

```{r, eval=TRUE, echo=FALSE}
sf_stages
```

The spatial data is represented as points at the moment.
Summarise points per stage, combining them into 'multipoints' and one row
per stage:

```{r, code=df_manifest$command[df_manifest$name == "sf_stages_multipoint"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(sf_stages_multipoint)
```

```{r, eval=TRUE, echo=FALSE}
sf_stages_multipoint
```

Cast into lines with this operation:

```{r sf_stages_line, code=df_manifest$command[df_manifest$name == "sf_stages_line"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(sf_stages_line)
```

```{r, eval=TRUE, echo=FALSE}
sf_stages_line
```

We can now plot the data using known 'tidyverse' [@cit_tidyverse] techniques.
To Include an underlying map, 'ggspatial' [@cit_ggspatial] is used.

```{r, file="R/vis_stages_line.R"}
```

```{r, code=df_manifest$command[df_manifest$name == "gg_stages_line"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(gg_stages_line)
```

```{r, eval=TRUE, echo=FALSE, preview=TRUE}
gg_stages_line
```

With 'leaflet' [@cit_leaflet]  we can also have an interactive look:

```{r, file="R/vis_stages_line_interactive.R"}
```

```{r, code=df_manifest$command[df_manifest$name == "gg_stages_line_interactive"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(gg_stages_line_interactive)
```

```{r, eval=TRUE, echo=FALSE}
gg_stages_line_interactive
```

# Conclusion

I think I found my perfect spot. By downloading the data, turning it into
a spatial format and plotting it interactively, I could easily explore the
route. I hope this helps you to find your spot as well! Hope to see you there :)